Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-01-31 manuscript/HIAs.qmd
try different validation plots

@@

# Make map to show where the three municipalities are
# First, get centroids
centroids <- muni3 |>
  st_centroid()

@@

km_distance <- centroids |>
  st_distance() |>
  max() |>
  set_units("km") |>
  drop_units() |>
  round()

round(max(latlong$Y)-min(latlong$Y), 1)

@@

```{r readInfraStructureIndex}
#| eval: false

@@

corrCheck |>
  as_tibble() |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  count(i_ADSV_fct, infrastructureIndex) |>
  View()

# A first look
corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(indicator~year)
```

@@

# Using a colour gradient emphasises the first colour (dark green).
# Lets try discrete colours, and merge som classes to simplify
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    .default = "Good"
  )) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)

@@

>>>>

```{r centroids}
#| eval: true
# Make map to show where the three municipalities are
# First, get centroids
centroids <- muni3 |>
  st_centroid()
```

@@

@@

km_distance <- centroids |>
  st_distance() |>
  max() |>
  set_units("km") |>
  drop_units() |>
  round()
#round(max(latlong$Y)-min(latlong$Y), 1)
```

@@

```{r readInfraStructureIndex}
#| eval: true

@@

#corrCheck |>
#  as_tibble() |>
#  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
#         i_alien_fct = floor(round(i_alien*10,2))/10) |>
#  count(i_ADSV_fct, infrastructureIndex) |>
#  View()

# A first look
corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(indicator~year)
```
```{r}
#| eval: false

# Try looking across all years
# A first look
corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(~indicator)
```

@@

# Using a colour gradient emphasises the first colour (dark green).
# Lets try discrete colours, and merge som classes to simplify
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    indicatorValue <0.91 ~"Good",
    .default = "Perfect"
  ),
  condition = fct_reorder(condition, indicatorValue)
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  #scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)

```
```{r}
#| eval: false

# The figure changes a lot depending on which value gets the darker colour
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    indicatorValue <0.91 ~"Good",
    .default = "Perfect"
  ),
  condition = fct_reorder(condition, -indicatorValue)
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  #scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)
```

# If I look at this on a log scale, it also changes and becomes less convincing
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    indicatorValue <0.91 ~"Good",
    .default = "Perfect"
  ),
  condition = fct_reorder(condition, indicatorValue)
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)

@@

# HIAs 1 and 2 are quite similar, so perhaps I can combine them
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    indicatorValue <0.91 ~"Good",
    .default = "Perfect"
  ),
  condition = fct_reorder(condition, indicatorValue),
  HIA = case_when(
    infrastructureIndex %in% c(1,2) ~ "1 & 2",
    .default = factor(infrastructureIndex)
  )
  ) |>
  ggplot(aes(
    x = HIA, 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)
```
